---
title: "Assignment: Prediction Assignment Writeup"
author: "Antonio Maurandi López"
date: "7 de junio de 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, echo=T, results="asis", cahe=TRUE)
options(scipen = 1, digits = 3)  # set default digits

library(pander)
library(caret)

# library(xtable)
# library("lattice")
```

Using devices such as _Jawbone Up_, _Nike FuelBand_, and _Fitbit_ it is now possible to collect a large amount of data about personal activity relatively inexpensively. 

Our goal is to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. 


# Loading and processing the data


Exploration of the data and data structure.

```{r, cache=TRUE}

# download data from the source url
filepath     <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
filepath     <- "pml-training.csv"
dftr <- read.table(filepath,  sep="," , header=T)
filepath     <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
filepath     <- "pml-testing.csv"
dfte <- read.table(filepath,  sep="," , header=T)
```

## Missing values: `NA` 

There are variables with a lot of `NA` values, we will use only those variables which are not `NA` always, let set the criteria at 80% of data available, not `NA`.

```{r}
n         <- nrow(dftr)
f.numofna <- function (vector){
    return( sum(is.na (vector)))
}

x       <- as.vector(apply(dftr, 2, f.numofna)/n)
table(x)
varsnona <- (x<0.8)


jj<- dftr[,varsnona]

kk<-lapply(jj,data.class)
# pander::pander(kk)
pander(kk)
dftr2 <- dftr[,varsnona]
n1<- length(names(dftr))
n2<- length(names(dftr2))
```

We will keep `r n2` variables instead of the original `r n1` variables.


It is interesting to check if there are any class more presnet tahn others
```{r}
tt<-table(dftr2$classe)
pander::pander(tt,caption="absolute frecuency of classes of variebl classe")
tt<- prop.table(tt)*100
pander::pander(tt,caption="percentage table")

```

We can see that there are 5 different clasifications and that all of them are arround 16% and 29%


# Classification


# Trainning and prove data sets


We will use a 60% training set, 40% prove set of the total data set with clsification (`classe`).

```{r}
set.seed(pi)
indices  <- createDataPartition(dftr2$classe, p=0.6)

df1 <- dftr2[indices[[1]],] # trainning data set
df2 <- dftr2[-indices[[1]],] # proving data set
```



## Preprocess


```{r}
zero.var   <- nearZeroVar(df1, saveMetrics=TRUE)
pander(zero.var)
varswithnonzv <- row.names(zero.var[zero.var$nzv==FALSE,])
df1<- df1[,varswithnonzv]
# nrow(df1)
# ncol(df1)
# as.data.frame(names(df1))
classe<- df1$classe
df1<-df1[,-c(2,5)] # quito dos factotes : nombre y time no se q
df1$classe<-NULL   # guardamos la var clasificacion (outcome) en otra yu la quitamos del cto de datos
```

Remove candidate variables with _near zero variance_. 
Now all variables have enough variance.


```{r}
require(corrplot)
corrplot.mixed(cor(df1), lower="circle", upper="color", 
               tl.pos="lt", diag="n", order="hclust", hclust.method="complete")
```

```{r, eval=FALSE}
featurePlot(df1, classe,  "ellipse")# sale una kk
```


```{r, eval=FALSE}

# esto es tela, se come el pc por los pies
# t-Distributed Stochastic Neighbor Embedding
tsne = Rtsne(as.matrix(df1), check_duplicates=FALSE, pca=TRUE, 
              perplexity=30, theta=0.5, dims=2)
embedding = as.data.frame(tsne$Y)
embedding$Class = class
g = ggplot(embedding, aes(x=V1, y=V2, color=Class)) +
  geom_point(size=1.25) +
  guides(colour=guide_legend(override.aes=list(size=6))) +
  xlab("") + ylab("") +
  ggtitle("t-SNE 2D Embedding of 'Classe' Outcome") +
  theme_light(base_size=20) +
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank())
print(g)
```


# `Ramndom Forest`

```{r, cahe=TRUE, eval=FALSE}
library(caret)
set.seed(pi)
fit.rf <-train(classe~., data=dftr2, method="rf")



library(ElemStatLearn)
data(vowel.train)
data(vowel.test) 
set.seed(33833)
vowel.test$y <- factor(vowel.test$y)
vowel.train$y <- factor(vowel.train$y)
suppressMessages(library(caret))
rfmodel <- suppressMessages(train(y~., data=vowel.train, method="rf"))
```


```{r, eval=FALSE}
# gbmmodel <- suppressMessages(train(y~., data=vowel.train, method="gbm"))

rf.pred <- predict(fit.rf, dftest[,varsnona])
# gbm.result <- predict(gbmmodel, vowel.test)

confusionMatrix(dftrai2$classe, rf.pred)$overall['Accuracy']
```


