---
title: "Assignment: Prediction Assignment Writeup"
author: "Antonio Maurandi López"
date: "21 de junio de 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, echo=T, results="asis", cahe=TRUE)
options(scipen = 1, digits = 3)  # set default digits

library(pander)
library(caret)

library(devtools)
# install_github("ujjwalkarn/xda")
library(xda)

# library(xtable)
# library("lattice")
```

Using devices such as _Jawbone Up_, _Nike FuelBand_, and _Fitbit_ it is now possible to collect a large amount of data about personal activity relatively inexpensively. 
Our goal is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. 
We will use data  data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants to predict  the "classe" variable.

# Loading and processing the data


Exploration of the data.

```{r, cache=TRUE}

# download data from the source url
filepath     <- "XXXXX"
filepath     <- "pml-training.csv"
train <- read.table(filepath,  sep="," , header=T)
filepath     <- "XXX"
filepath     <- "pml-testing.csv"
test <- read.table(filepath,  sep="," , header=T)
```

## Missing values: `NA` 

There are variables with a lot of `NA` values, we will use only those variables which are not `NA` always, let set the criteria at 80% of data available, not `NA`.

```{r}
n         <- nrow(train)
f.numofna <- function (vector){
    return( sum(is.na (vector)))
}

x       <- as.vector(apply(train, 2, f.numofna)/n)
pander(table(x), caption = "Number of variables with a percentaje of missiing values: `0%` or `97%`")
varsnona <- (x<0.8)

n1    <- length(names(train))
train <- train[,varsnona]
test  <- test[,varsnona]

# kk<-lapply(train,data.class)
# pander(kk)
n2 <- length(names(train))
```

We will keep `r n2` variables instead of the original `r n1` variables.

## Outcome variable `classe`

It is interesting to check if there are any class more presnet tahn others
```{r}
tt<-table(train$classe)
pander::pander(tt,caption="absolute frecuency of classes of variebl classe")
tt<- prop.table(tt)*100
pander::pander(tt,caption="percentage table")

```

We can see that there are 5 different clasifications and that all of them are arround 16% and 29% of the total data available.


```{r, results="asis"}
pander(numSummary(train)[,-c(7:17)], caption="Preliminary descriptives of numerical variables in the trainning dataset.")
```


```{r, results="asis"}
pander(charSummary(train), caption="Preliminary descriptives of non numerical variables in the trainning dataset.")
```


## Standaritation

It's interestilng to _scale_ and _center_ the data  it to  get models less influenced by the diferent scale of the predictor variables.

```{r}
procValues <- preProcess(train, method = c("center", "scale"))
trainN     <- predict(procValues, train)
testN      <- predict(procValues, test)
```


We will work with the data set witout the classification variable `classe`.


aml! poor aqui voy
```{r}
myclasse     <- trainN$classe  # classification/outcome variable
trainN       <- subset(trainN, select = -c(classe))    
testN        <- subset(testN , select = -c(problem_id)) 
```

Now we have a trainning data set with `r ncol(trainN)` variables and `r nrow(trainN)` observations.


# Model

## Reduction of dimensionality

By means of PCA we will consider a problem with a lower dimensionality

```{r, eval=FALSE}
names(trainN)
precPCA     <- preProcess(trainN, method = "pca",thresh = 0.9)
trainN_PCA  <- predict(precPCA,  trainN )
str(trainN_PCA)
testN_PCA   <- predict(precPCA,  testN )
```


## Cross validation

We will use a 60% training set, 40% prove set of the total data set with clsification (`classe`).

```{r}
set.seed(pi)
casostest1  <- createDataPartition(myclasse, p=0.6, list = FALSE)
str(casostest1)

train1      <- trainN [casostest1,]  # trainning data set
train2      <- trainN [-casostest1,] # proving data set
```


We will use to build a model a trainning data set of `r nrow(train1)` observations that is the 60% of the data in the origibal trainning dataset. We will test the model in a test data set of `r nrow(train2)`  that is the remaining 40% of the data of the originla trainning dataset.








```{r}
df<- cbind(train1, myclasse[casostest1])

method <- "rf"
system.time(trainingModel <- train(df$myclasse ~ ., data=df, method="rf"))
```






This are the outputs of the model:

```{r, eval=FALSE}
load("my_model2.rda")
rf.final
```

We see that the best `Accuracy` is obtained for mtry = `rf.final$bestTune`, so this one is the chosen one.







## Preprocess


```{r, eval=FALSE}
zero.var   <- nearZeroVar(df1, saveMetrics=TRUE)
pander(zero.var)
varswithnonzv <- row.names(zero.var[zero.var$nzv==FALSE,])
df1<- df1[,varswithnonzv]
# nrow(df1)
# ncol(df1)
# as.data.frame(names(df1))
classe<- df1$classe
df1<-df1[,-c(2,5)] # quito dos factotes : nombre y time no se q
df1$classe<-NULL   # guardamos la var clasificacion (outcome) en otra yu la quitamos del cto de datos
```

Remove candidate variables with _near zero variance_. 
Now all variables have enough variance.




# `Ramndom Forest`

```{r, cahe=TRUE, eval=FALSE}
library(caret)
set.seed(pi)
fit.rf <-train(classe~., data=dftrai2, method="rf")



library(ElemStatLearn)
data(vowel.train)
data(vowel.test) 
set.seed(33833)
vowel.test$y <- factor(vowel.test$y)
vowel.train$y <- factor(vowel.train$y)
suppressMessages(library(caret))
rfmodel <- suppressMessages(train(y~., data=vowel.train, method="rf"))
```


```{r, eval=FALSE}
# gbmmodel <- suppressMessages(train(y~., data=vowel.train, method="gbm"))

rf.pred <- predict(fit.rf, dftest[,varsnona])
# gbm.result <- predict(gbmmodel, vowel.test)

confusionMatrix(dftrai2$classe, rf.pred)$overall['Accuracy']
```



